---
title: 'Reproducible Research: Peer Assessment 1'
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
1. We load the data
2. Convert the date into correct date format
3. Mutate the weekday/weekend column with each row
```{r loadProcessData, echo=TRUE}
activity <- read.csv("./activity.csv")
library(dplyr)
library(lubridate)
activity$date <- as.Date(activity$date)
activity <- mutate(activity,isWeekend = ifelse((wday(activity$date) == 7 | wday(activity$date) == 1), "Weekend", "Weekday"))
```



## What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day
2. Plot the histogram
3. Calculate the mean and median of total number of steps per day
```{r meanSteps, echo=TRUE}
tSteps <- activity %>% group_by(date) %>% summarize(totalStepsPerDay = sum(steps),na.rm=TRUE)
hist(tSteps$totalStepsPerDay,xlab = "Total Steps Per Day", main = "Histogram of Total number of Steps Per Day")
tMean <- mean(tSteps$totalStepsPerDay,na.rm = TRUE)
tMedian <- median(tSteps$totalStepsPerDay, na.rm = TRUE)
tMean <- format(round(tMean,2),nsmall = 2)
```
The mean total number of steps taken per day is `r tMean` and median is `r tMedian`

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
